---
title: Supplementary information to report "How to measure response diversity"
author: "Samuel R.P-J. Ross1, Owen L. Petchey, Takehiro Sasaki, David W. Armitage"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      echo = FALSE)
```

```{r}
rm(list = ls())
library(tidyverse)
library(patchwork)
library(ggtext)
library(mgcv)
library(gratia)
library(hypervolume)
library(here)

source(here("r/our_functions.r"))

```

# Introduction

To be written.

# SD - Two dimensions of response diversity

I created communities containing species, and gave each species $i$ a value for the first derivative of it's response to an environmental variable ($x_i$).

I think there are two dimensions (at least) of the diversity of these values. First, the spread of the distribution, which is what we already use in the ms. Here I use the standard deviation, i.e., sd($x$). This is not affected by the mean of $x$, i.e. is not affect by if the values of $x$ span zero or do not span zero. (I use sd($x$) rather than CV($x$) because the CV is very much affected by values of mean($x$) close to zero.)

We do, however, (I think) consider it important if the values of $x$ include some negative and some positive values, and therefore should quantify this. I have done so with (sd($x$) - sd($|x|$)) / sd($x$). (I came up with this just by thinking... probably should have done some reading also.) This varies from zero to one, being zero when all $x$ have the same sign, and being one when there are only two unique values of $x$, say $x_1$ and $x_2$ and $x_1 = -x_2$. I'm  sure this is **not** the best thing to do, or even if it is very well justified, but it is at least independent of sd($x$), as a measure of this dimension should be, as is shown below in some examples.


```{r}
## Create the communities
num_spp <- c(2, 4, 8, 16)
max_minus_min <- round(seq(1, 15, length = 45),3)
mean <- round(seq(-10, 10, length = 49),3)
rd_vals <- function(num_spp, max_minus_min, mean) {
  seq(mean-max_minus_min/2, mean+max_minus_min/2, length = num_spp)
}
dd <- crossing(num_spp,
               max_minus_min,
               mean) %>%
  rowwise() %>%
  mutate(rd_vals = list(rd_vals(num_spp, max_minus_min, mean))) %>%
  unnest(cols = rd_vals)
```


```{r}
## Set which measure of response diversity used
rd_stat <- "sd"
```


```{r}
## calculate the diversity metrics
summary_stats <- dd %>%
  group_by(num_spp, max_minus_min, mean) %>%
  summarise(sd = resp_div(rd_vals, stat = rd_stat, sign_sens = FALSE),
            #cv = sd/mean(rd_vals),
            #cv_abs = sd/abs(mean(rd_vals)),
            #sd_abs_vals = sd(abs(rd_vals)),
            sign_sensitive = resp_div(rd_vals, stat = rd_stat, sign_sens = TRUE)) %>%
  pivot_longer(names_to = "variable", values_to = "value", 4:5)

## wrangle the data for easier plotting
all_data <- dd %>%
  mutate(variable = "first_deriv_vals",
         value = rd_vals) %>%
  select(-rd_vals) %>%
  bind_rows(summary_stats) %>%
  mutate(variable = as_factor(variable))
```

## Communities varying only in spread

Here we look at communities that only vary in the spread of $x$ and that do not differ in the second dimension (overlap with zero). As expected, sd($x$) varies, but the measure of the overlap with zero does not. In the first example, all communities have equal overlap with zero (Fig. \@ref(fig:XXX1)) and in the second none of the communities overlap zero (Fig. \@ref(fig:XXX2)).

(ref:XXX1) Each column of dots corresponds to one community. Top panel shows the values of the first derivative of each of the species in the community. Middle is the sd($x$): the spread of the distribution. Bottom panel is the (sd($x$) - sd($|x|$)) / sd($x$) -- a measure of overlap with zero.

```{r XXX1,  include=TRUE, fig.cap='(ref:XXX1)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 0
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community") 



```

(ref:XXX2) Each column of dots corresponds to one community. Top panel shows the values of the first derivative of each of the species in the community. Middle is the sd($x$): the spread of the distribution. Bottom panel is the (sd($x$) - sd($|x|$)) / sd($x$) -- a measure of overlap with zero.

```{r XXX2,  include=TRUE, fig.cap='(ref:XXX2)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 2.5
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp,
         max_minus_min < 5) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community")
```


## Communities varying only in overlap

Here we look at communities that only vary in the overlap of $x$ with zero, and that do not differ in spread. As expected, sd($x$) does not vary, but the measure of the overlap with zero does. It is zero while there is not overlap, then increases to a maximum when the mean is zero, and then decreases again (Fig. \@ref(fig:XXX3).

(ref:XXX3) Each column of dots corresponds to one community. Top panel shows the values of the first derivative of each of the species in the community. Middle is the sd($x$): the spread of the distribution. Bottom panel is the (sd($x$) - sd($|x|$)) / sd($x$) -- a measure of overlap with zero.

```{r XXX3,  include=TRUE, fig.cap='(ref:XXX3)', fig.align="center", fig.height=4, fig.width=4}
this_max_minus_min <- 5.455
this_num_spp <- 8
all_data %>%
  filter(max_minus_min == this_max_minus_min,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = mean, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Mean of first derivative values in community") 
```




## Communities varying in both spread and overlap

Here we look at communities that vary in both spread and the overlap. As expected, sd($x$) increases but the overlap measure does not. Then the overlap measure increases as the amount of overlap with zero increases (Fig. \@ref(fig:XXX4).

(ref:XXX4) Each column of dots corresponds to one community. Top panel shows the values of the first derivative of each of the species in the community. Middle is the sd($x$): the spread of the distribution. Bottom panel is the (sd($x$) - sd($|x|$)) / sd($x$) -- a measure of overlap with zero. The data at values of less than 5 on the x-axis are the same as the data in Fig. \@ref(fig:XXX2).

```{r XXX4,  include=TRUE, fig.cap='(ref:XXX4)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 2.5
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community")
```

## Varying number of species

This is the one I'm really not satisfied with. I think that either, or perhaps even both, dimension of response diversity should increase with species richness, but both actually decrease (Fig. \@ref(fig:XXX5)). I think this is why Kevin Gaston and I opted for the total branch length of the dendrogram of hierarchical clustering of the distance matrix -- it has this lovely property of only increasing (or staying the same) with increasing number of species. Anyway, I hand this over to you for contemplation, before continuing.

(ref:XXX5) Each column of dots corresponds to one community. Top panel shows the values of the first derivative of each of the species in the community. Middle is the sd($x$): the spread of the distribution. Bottom panel is the (sd($x$) - sd($|x|$)) / sd($x$) -- a measure of overlap with zero.

```{r XXX5,  include=TRUE, fig.cap='(ref:XXX5)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 0
this_num_spp <- 8
this_max_minus_min <- 5.455
all_data %>%
  filter(mean == this_mean,
         max_minus_min == this_max_minus_min) %>%
  ggplot(aes(x = num_spp, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Number of species") 



```

# Range -- Two dimensions of response diversity

This section duplicates the previous one, but with range instead of SD.
Not that text has not yet been revise to read range rather than SD.

```{r}
## Set which measure of response diversity used
rd_stat <- "range"
```

```{r}
## calculate the diversity metrics
summary_stats <- dd %>%
  group_by(num_spp, max_minus_min, mean) %>%
  summarise(range = resp_div(rd_vals, stat = rd_stat, sign_sens = FALSE),
            #cv = sd/mean(rd_vals),
            #cv_abs = sd/abs(mean(rd_vals)),
            #sd_abs_vals = sd(abs(rd_vals)),
            sign_sensitive = resp_div(rd_vals, stat = rd_stat, sign_sens = TRUE)) %>%
  pivot_longer(names_to = "variable", values_to = "value", 4:5)

## wrangle the data for easier plotting
all_data <- dd %>%
  mutate(variable = "first_deriv_vals",
         value = rd_vals) %>%
  select(-rd_vals) %>%
  bind_rows(summary_stats) %>%
  mutate(variable = as_factor(variable))
```

## Communities varying only in spread



(ref:XXX6) XXX

```{r XXX6,  include=TRUE, fig.cap='(ref:XXX6)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 0
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community") 



```

(ref:XXX7) XXX

```{r XXX7,  include=TRUE, fig.cap='(ref:XXX7)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 2.5
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp,
         max_minus_min < 5) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community")
```


## Communities varying only in overlap

Here we look at communities that only vary in the overlap of $x$ with zero, and that do not differ in spread. As expected, sd($x$) does not vary, but the measure of the overlap with zero does. It is zero while there is not overlap, then increases to a maximum when the mean is zero, and then decreases again (Fig. \@ref(fig:XXX3).

(ref:XXX8) XXX

```{r XXX8,  include=TRUE, fig.cap='(ref:XXX8)', fig.align="center", fig.height=4, fig.width=4}
this_max_minus_min <- 5.455
this_num_spp <- 8
all_data %>%
  filter(max_minus_min == this_max_minus_min,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = mean, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Mean of first derivative values in community") 
```




## Communities varying in both spread and overlap

(ref:XXX9) XXX

```{r XXX9,  include=TRUE, fig.cap='(ref:XXX9)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 2.5
this_num_spp <- 8
all_data %>%
  filter(mean == this_mean,
         num_spp == this_num_spp) %>%
  ggplot(aes(x = max_minus_min, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Range of first derivative values in community")
```

## Varying number of species

(ref:XXX10) XXX

```{r XXX10,  include=TRUE, fig.cap='(ref:XXX10)', fig.align="center", fig.height=4, fig.width=4}
this_mean <- 0
this_num_spp <- 8
this_max_minus_min <- 5.455
all_data %>%
  filter(mean == this_mean,
         max_minus_min == this_max_minus_min) %>%
  ggplot(aes(x = num_spp, y = value)) +
  geom_point() +
  geom_hline(yintercept = 0, lty="dashed") +
  facet_wrap( ~ variable, scales = "free_y", ncol = 1) +
  xlab("Number of species") 
```

# Producing Figures 2-3 based on simulated data

This section will produce Figures 2-3 from the manuscript. The figures are inspired by a range of examples from the literature (see main text), but are based on simulated data.

First, we'll generate a series of simulated communities that have high, medium, or low response diversity to an environmental variable [x], based on correlated or uncorrelated responses, and responding either linearly or nonlinearly.

```{r}
# generate an environmental variable (x), as a simple sequence of numbers
x = seq(0, 1, length.out = 100)

# generate species with different responses to the environment 
# Linear responses:
y1a <- 0.8-1*x; y2a <- 0.75-1* x; y3a <-0.85-1*x # responses are similar in direction and magnitude (low response diversity)
y1b <- 0.5-0.2*x; y2b <- 0.75-1*x; y3b <-1-2.5*x # responses are similar in direction but not magnitude (medium response diversity)
y1c <- 0.6-0.75*x; y2c <- rev(y1c); y3c <- rep(0.35, length(x)) # responses are not similar in direction or magnitude (high response diversity)

# Nonlinear responses:
y1d <- range01(dnorm(x,0.5,0.3)); y2d <- range01(dnorm(x,0.475,0.3)); y3d <- range01(dnorm(x,0.525,0.3)) # responses are similar in direction and magnitude (low response diversity)
y1e <- range01(dnorm(x,0.5,0.3)); y2e <- range01(dnorm(x,0.2,0.3)); y3e <- range01(dnorm(x,0.8,0.3)) # responses are similar in direction but not magnitude (medium response diversity)
y1f <- exp(-3*x); y2f <- rev(y1f); y3f <- 1-(y1f+y2f) # responses are not similar in direction or magnitude (high response diversity)
```

We'll model these species-environment relationships using generalized additive models (GAMs).

```{r}
# model individual species~environment relationships using GAM
m1a <- gam(y1a~s(x)); m2a <- gam(y2a~s(x)); m3a <- gam(y3a~s(x))
m1b <- gam(y1b~s(x)); m2b <- gam(y2b~s(x)); m3b <- gam(y3b~s(x))
m1c <- gam(y1c~s(x)); m2c <- gam(y2c~s(x)); m3c <- gam(y3c~s(x))
m1d <- gam(y1d~s(x)); m2d <- gam(y2d~s(x)); m3d <- gam(y3d~s(x))
m1e <- gam(y1e~s(x)); m2e <- gam(y2e~s(x)); m3e <- gam(y3e~s(x))
m1f <- gam(y1f~s(x)); m2f <- gam(y2f~s(x)); m3f <- gam(y3f~s(x))
```

Then we calculate the first derivatives of these relationships.

```{r}
# calculate first derivatives from models
d1a <- derivatives(m1a); d2a <- derivatives(m2a); d3a <- derivatives(m3a)
d1b <- derivatives(m1b); d2b <- derivatives(m2b); d3b <- derivatives(m3b)
d1c <- derivatives(m1c); d2c <- derivatives(m2c); d3c <- derivatives(m3c)
d1d <- derivatives(m1d); d2d <- derivatives(m2d); d3d <- derivatives(m3d)
d1e <- derivatives(m1e); d2e <- derivatives(m2e); d3e <- derivatives(m3e)
d1f <- derivatives(m1f); d2f <- derivatives(m2f); d3f <- derivatives(m3f)
```

Then we'll measure response diversity using the approach above. 

```{r}

# calculate response diversity for model a
rdiv_a<-data.frame('data' = d1a$data, 'sp.1' = d1a$derivative, 'sp.2' = d2a$derivative, 'sp.3' = d3a$derivative)
rdiv_a$rdiv<-apply(rdiv_a[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_a$sign<-apply(rdiv_a[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_a$Med<-median(rdiv_a$rdiv)

# calculate response diversity for model b
rdiv_b<-data.frame('data' = d1b$data, 'sp.1' = d1b$derivative, 'sp.2' = d2b$derivative, 'sp.3' = d3b$derivative)
rdiv_b$rdiv<-apply(rdiv_b[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_b$sign<-apply(rdiv_b[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_b$Med<-median(rdiv_b$rdiv)

# calculate response diversity for model c
rdiv_c<-data.frame('data' = d1c$data, 'sp.1' = d1c$derivative, 'sp.2' = d2c$derivative, 'sp.3' = d3c$derivative)
rdiv_c$rdiv<-apply(rdiv_c[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_c$sign<-apply(rdiv_c[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_c$Med<-median(rdiv_c$rdiv)

# calculate response diversity for model d
rdiv_d<-data.frame('data' = d1d$data, 'sp.1' = d1d$derivative, 'sp.2' = d2d$derivative, 'sp.3' = d3d$derivative)
rdiv_d$rdiv<-apply(rdiv_d[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_d$sign<-apply(rdiv_d[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_d$Med<-median(rdiv_d$rdiv)

# calculate response diversity for model e
rdiv_e<-data.frame('data' = d1e$data, 'sp.1' = d1e$derivative, 'sp.2' = d2e$derivative, 'sp.3' = d3e$derivative)
rdiv_e$rdiv<-apply(rdiv_e[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_e$sign<-apply(rdiv_e[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_e$Med<-median(rdiv_e$rdiv)

# calculate response diversity for model f
rdiv_f<-data.frame('data' = d1f$data, 'sp.1' = d1f$derivative, 'sp.2' = d2f$derivative, 'sp.3' = d3f$derivative)
rdiv_f$rdiv<-apply(rdiv_f[,c(2:4)], 1, resp_div, stat = "range", sign_sens = F)
rdiv_f$sign<-apply(rdiv_f[,c(2:4)], 1, resp_div, stat = "range", sign_sens = T)
rdiv_f$Med<-median(rdiv_f$rdiv)

```

Now we have all the data required to produce the figures from the manuscript. First, we'll produce *Figure 2*, which looks at linear performance-environment relationships: 

```{r}
# Y value range
ymax<-max(y1a,y2a,y3a,y1b,y2b,y3b,y1c,y2c,y3c);ymin<-min(y1a,y2a,y3a,y1b,y2b,y3b,y1c,y2c,y3c)
dmax<-max(d1a$derivative,d2a$derivative,d3a$derivative,d1b$derivative,d2b$derivative,d3b$derivative,d1c$derivative,d2c$derivative,d3c$derivative)
dmin<-min(d1a$derivative,d2a$derivative,d3a$derivative,d1b$derivative,d2b$derivative,d3b$derivative,d1c$derivative,d2c$derivative,d3c$derivative)
rmax<-max(rdiv_a$rdiv,rdiv_b$rdiv,rdiv_c$rdiv);rmin<-0
smax<-1;smin<-0

# panel a
Fig_2a<-ggplot(data = NULL,mapping = aes(x = x,y = y1a)) +
  theme_classic(base_size = 14) + 
  labs(x = "Spring frost",y = "Chlorophyll production",title = "Low response diversity",tag = "a)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2a),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3a),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel b
Fig_2b<-ggplot(data = NULL,mapping = aes(x = x,y = y1b)) +
  theme_classic(base_size = 14) + 
  labs(x = "Urbanisation",y = "Macroinvertebrate abundance",title = "Medium response diversity",tag = "b)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2b),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3b),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel c
Fig_2c<-ggplot(data = NULL,mapping = aes(x = x,y = y1c)) +
  theme_classic(base_size = 14) + 
  labs(x = "Winter temperature",y = "Bee mass loss",title = "High response diversity",tag = "c)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2c),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3c),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel d
Fig_2d<-ggplot(data = NULL,mapping = aes(x = d1a$data,y = d1a$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Spring frost",y = "Derivative (chlorophyll product.)",tag = "d)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2a$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3a$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel e
Fig_2e<-ggplot(data = NULL,mapping = aes(x = d1b$data,y = d1b$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Urbanisation",y = "Derivative (macroinvert. abundance)",tag = "e)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2b$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3b$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel f
Fig_2f<-ggplot(data = NULL,mapping = aes(x = d1c$data,y = d1c$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Winter temperature",y = "Derivative (bee mass loss)",tag = "f)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2c$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3c$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel g
Fig_2ga<-ggplot(data = rdiv_a,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "g)") + 
  geom_hline(yintercept = rdiv_a$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_a$Med+0.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_a$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_2gb<-ggplot(data = rdiv_a,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Spring frost",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_2gb<-Fig_2gb + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# panel h
Fig_2ha<-ggplot(data = rdiv_b,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "h)") + 
  geom_hline(yintercept = rdiv_b$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_b$Med-0.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_b$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_2hb<-ggplot(data = rdiv_b,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Urbanisation",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_2hb<-Fig_2hb + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# panel i
Fig_2ia<-ggplot(data = rdiv_c,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "i)") + 
  geom_hline(yintercept = rdiv_c$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_c$Med+0.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_c$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_2ib<-ggplot(data = rdiv_c,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Winter temperature",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_2ib<-Fig_2ib + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

(Fig_2a + Fig_2b + Fig_2c)/(Fig_2d + Fig_2e + Fig_2f)/(Fig_2ga + Fig_2ha + Fig_2ia)/(Fig_2gb + Fig_2hb + Fig_2ib) + plot_layout(heights = c(3,3,3,1))
```

Next, we'll produce *Figure 3* to look at _nonlinear_ performance-environment relationships: 

```{r}
# Y value range
ymax<-max(y1d,y2d,y3d,y1e,y2e,y3e,y1f,y2f,y3f);ymin<-min(y1d,y2d,y3d,y1e,y2e,y3e,y1f,y2f,y3f)
dmax<-max(d1d$derivative,d2d$derivative,d3d$derivative,d1e$derivative,d2e$derivative,d3e$derivative,d1f$derivative,d2f$derivative,d3f$derivative)
dmin<-min(d1d$derivative,d2d$derivative,d3d$derivative,d1e$derivative,d2e$derivative,d3e$derivative,d1f$derivative,d2f$derivative,d3f$derivative)
rmax<-max(rdiv_d$rdiv,rdiv_e$rdiv,rdiv_f$rdiv);rmin<-0
smax<-1;smin<-0

# panel a
Fig_3a<-ggplot(data = NULL,mapping = aes(x = x,y = y1d)) +
  theme_classic(base_size = 14) + 
  labs(x = "Flowering plant abundance",y = "Flowering plant rate of increase",title = "Low response diversity",tag = "a)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2d),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3d),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel b
Fig_3b<-ggplot(data = NULL,mapping = aes(x = x,y = y1e)) +
  theme_classic(base_size = 14) + 
  labs(x = "Grazing pressure",y = "Seedling establishment",title = "Medium response diversity",tag = "b)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2e),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3e),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel c
Fig_3c<-ggplot(data = NULL,mapping = aes(x = x,y = y1f)) +
  theme_classic(base_size = 14) + 
  labs(x = "Fire frequency",y = "Tree density",title = "High response diversity",tag = "c)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = y2f),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = y3f),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(ymin,ymax))

# panel d
Fig_3d<-ggplot(data = NULL,mapping = aes(x = d1d$data,y = d1d$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Flowering plant abundance",y = "Derivative (rate of increase)",tag = "d)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2d$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3d$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel e
Fig_3e<-ggplot(data = NULL,mapping = aes(x = d1e$data,y = d1e$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Grazing pressure",y = "Derivative (seedling establish.)",tag = "e)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2e$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3e$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel f
Fig_3f<-ggplot(data = NULL,mapping = aes(x = d1f$data,y = d1f$derivative)) +
  theme_classic(base_size = 14) + 
  labs(x = "Fire frequency",y = "Derivative (tree density)",tag = "f)") + 
  geom_hline(yintercept = 0,
             lty=2) +
  geom_line(col = Isfahan1[[1]][2]) + 
  geom_line(aes(y = d2f$derivative),
            col = Isfahan1[[1]][4]) + 
  geom_line(aes(y = d3f$derivative),
            col = Isfahan1[[1]][6]) + 
  lims(y = c(dmin,dmax))

# panel g
Fig_3ga<-ggplot(data = rdiv_d,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "g)") + 
  geom_hline(yintercept = rdiv_d$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_d$Med+1.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_d$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_3gb<-ggplot(data = rdiv_d,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Flowering plant abundance",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_3gb<-Fig_3gb + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# panel h
Fig_3ha<-ggplot(data = rdiv_e,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "h)") + 
  geom_hline(yintercept = rdiv_e$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_e$Med+1.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_e$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_3hb<-ggplot(data = rdiv_e,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Grazing pressure",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_3hb<-Fig_3hb + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# panel i
Fig_3ia<-ggplot(data = rdiv_f,mapping = aes(x = data,y = rdiv)) +
  theme_classic(base_size = 14) + 
  labs(x = NULL,y = "Range (derivatives)",tag = "i)") + 
  geom_hline(yintercept = rdiv_f$Med,lty=2) +
  geom_line() + 
  geom_richtext(mapping = aes(x = 0.85,
                              y = rdiv_f$Med-1.25),
                size=4.5,
                label.color = NA,
                label = paste("RDiv^Med =",paste0(round(rdiv_f$Med,digits = 2)))) +
  lims(y = c(rmin,rmax))
Fig_3ib<-ggplot(data = rdiv_f,mapping = aes(x = data,y = sign)) +
  theme_bw(base_size = 14)+
  labs(x = "Fire frequency",y = "Sign") + 
  geom_line() + 
  lims(y = c(smin,smax))
Fig_3ib<-Fig_3ib + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

(Fig_3a + Fig_3b + Fig_3c)/(Fig_3d + Fig_3e + Fig_3f)/(Fig_3ga + Fig_3ha + Fig_3ia)/(Fig_3gb + Fig_3hb + Fig_3ib) + plot_layout(heights = c(3,3,3,1))

```

# Leary & Petchey (2009) reanalysis

This so far only reads in the r and K data and plots their response to temperature for the four species.

```{r}
#rm(list = ls())
r_vals <- read_csv(here("data/Leary_and_Petchey_r_values.csv")) %>%
  mutate(parameter = "r")
K_vals <- read_csv(here("data/Leary_and_Petchey_K_values.csv")) %>%
  mutate(x = log10(x), parameter = "log10_K")
rK_vals <- rbind(r_vals, K_vals) %>%
  rename(value = x)
rm(r_vals, K_vals)
```

```{r}
rK_vals %>%
ggplot(aes(x = temperature, y = value, col = species)) +
  geom_point() +
  geom_smooth() +
  facet_wrap( ~ parameter, scales = "free", nrow = 2) 
```


# Results of the systematic review

Here are the results from the systematic review conducted by Samuel RP-J Ross in December 2021-January 2022. Columns are as follows: 

*Date* = Date on which paper was included in review.
*WoS_search* = Search string for reproducing systematic review (NA indicates papers not returned in WoS results).
*Handled_by* = Intitials of author handling each paper during screening (SRP-JR = Samuel RP-J Ross).
*Included* = Binary, included (1) or not (0) as an empirical paper that measures response diversity after all screening.
*Authors* = Short form author list of each article. 
*Title* = Title of article.
*Journal* = Journal of article publication. 
*Year* = Year of publication. 
*Vol* = Volume of publication. 
*Issue* = Issue of publication. 
*Pg_start* = First page number of publication.
*Pg_end* = Last page number of publication.
*Article_ID* = Identifier of articles without page numbers.
*DOI* = Digital object identifier (DOI) of article. 
*Empirical_study* = Does the study use empirical data? 1 = yes, 0 = no, 0.5 = sort of (Theory papers etc.).
*Measures_RD* = Does the empirical (or theoretical) study measure response diversity? 1 = yes, 0 = no, 0.5 = sort of.
*RD_traits* = For papers that measure response diversity, do they measure functional response trait diversity? 0 = no, 0.5 = sort of, FDis = yes, using Functional Dispersion, RaoQ = yes, using Rao's Quadratic Entropy.
*RD_Int_term* = For papers that measure response diversity, do they measure an interaction term between species responding to an environmental variable? 0 = no, 0.5 = sort of, 1 = yes.
*RD_Other* = For papers that measure response diversity, do they take another approach (i.e. not traits or interaction terms)? 0 = no, 1 = yes. 
*Method_Description* = Short written description to help clarify method (see individual articles for more details).
*Notes* = Any remaining comments from SRP-JR to help justify inclusion/exclusion, or additional comments.

```{r}
review <- read_csv(here("data/WoS_results.csv"))
review
```

